A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-11-06 imaging_demo/ms.tex
Update section 2 with suggested changes
From b7d9d50 (review):
=====================

In ms.tex,

(*) Update section 2 with suggested changes

Interferometers measure {\it visibilities} -- the degree of coherence between
electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
A visibility, $\widetilde{V}_p$, can be written as:
\begin{align}
  \widetilde{V}_p = &\left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle_t \label{eqn:cc-vis}\hfill\\
  		= &\left\langle \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a \right] \right. \nonumber\\
		&\qquad \times \left.\left[ \int {W}^\star_b(\hat{\mathbf{s'}})\,E^\star (\hat{\mathbf{s'}})\,e^{i2\pi\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}}\,\dif^2\hat{\mathbf{s'}} + \widetilde{n}^\star_b \right] \right\rangle_t \\
                  =& \iint {W}_a(\hat{\mathbf{s}}) {W}^\star_b(\hat{\mathbf{s'}}) \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t e^{-i2\pi(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})}\,\dif^2\hat{\mathbf{s}}\,\dif^2\hat{\mathbf{s'}},
\end{align}
where we have brought the time average into the integral under the assumption that 
the aperture illumination pattern does not change over the time-scale of the 
averaging and $\star$ denotes complex conjugation. This expression can be further 
simplified with the sky brightness, 
$I(\hat{\mathbf{s}})= \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) 
\right\rangle_t  \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the 
antenna pair sky power response function (or the directional antenna power pattern), 
$B_p(\hat{\mathbf{s}})~\equiv~{W}_a(\hat{\mathbf{s}})~{W}^\star_b(\hat{\mathbf{s}})$. 
The result is the visibility expressed in terms of the sky brightness, the power 
response, and uncorrelated noise terms which we group into $\widetilde{n}_p$.
\begin{equation}
\widetilde{V}_p = \int e^{-i2\pi\mathbf{u}_p\!\cdot\,\hat{\mathbf{s}}}\,B_p(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_p,
\end{equation}
where, the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the 
vector separation between the two antennas. This signifies that the visibility 
($\widetilde{V}_p$) measured between a
pair of antennas ($p$) is obtained by the multiplying the sky brightness
$I(\hat{\mathbf{s}})$ by the antenna power response $B_p(\hat{\mathbf{s}})$ and
Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
coordinates, which are then sampled at the locations of the antenna spacings (or
baselines), namely, $\mathbf{u}_p$, and added to the receiver noise $n_p$.

This can be equivalently re-written as:
\begin{align}\label{eqn:software-holography}
  \widetilde{V}_p &= \int \widetilde{B}(\mathbf{u}^\prime-\mathbf{u}) \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int} + n_p,
\end{align}
where, $\widetilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
convolution with $\widetilde{B}(\mathbf{u})$ in $uv$-space. This is the software
holographic equivalent of traditional FX correlator output.

@@

The intermediate step of gridding with the antenna power response can be 
expressed as a convolution of a data vector generated by gridding the electric fields 
directly with the antenna illumination pattern.
\begin{align}
&\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v})=\nonumber\\
&\qquad\left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]
\ast\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
\end{align}

We can then use the multiplication-convolution theorem to move the convolution in 
Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in Equation 
\ref{eqn:dirty-image-FX}.
\begin{align}
  \mathbf{I}^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}^{-1}(\mathbf{a},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a})\,\right|^2\right\rangle_t. \label{eqn:dirty-image-MOFF}
\end{align}
The term inside the angular brackets before squaring has a very similar form as
that in equation~\ref{eqn:dirty-image-FX}. It signifies that the measured antenna
electric fields are weighted by the antenna noise, weighted and gridded by the
antenna aperture kernel, Fourier transformed and finally squared to obtain the
same image estimated that would have been obtained using equation~
\ref{eqn:dirty-image-FX}.

>>>>

Interferometers measure {\it visibilities} -- the degree of coherence between
electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
A visibility, $\widetilde{V}_p$, can be written as:
\begin{align}
  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle_t \label{eqn:cc-vis}\hfill\\
  		&= \left\langle \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a \right] \right. \nonumber\\
		&\qquad \times \left.\left[ \int {W}^\star_b(\hat{\mathbf{s'}})\,E^\star (\hat{\mathbf{s'}})\,e^{i2\pi\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}}\,\dif^2\hat{\mathbf{s'}} + \widetilde{n}^\star_b \right] \right\rangle_t \\
                  &= \iint {W}_a(\hat{\mathbf{s}}) {W}^\star_b(\hat{\mathbf{s'}}) \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t e^{-i2\pi(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})}\,\dif^2\hat{\mathbf{s}}\,\dif^2\hat{\mathbf{s'}},
\end{align}
where we have brought the time average into the integral under the assumption that 
the aperture illumination pattern does not change over the time-scale of the 
averaging and $\star$ denotes complex conjugation. This expression can be further 
simplified with the sky brightness, 
$I(\hat{\mathbf{s}})= \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) 
\right\rangle_t  \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the 
antenna pair sky power response function (or the directional antenna power pattern), 
$B_p(\hat{\mathbf{s}})~\equiv~{W}_a(\hat{\mathbf{s}})~{W}^\star_b(\hat{\mathbf{s}})$. 
The result is the visibility expressed in terms of the sky brightness, the power 
response, and uncorrelated noise terms which we group into $\widetilde{n}_p$.
\begin{equation}
\widetilde{V}_p = \int e^{-i2\pi\mathbf{u}_p\!\cdot\,\hat{\mathbf{s}}}\,B_p(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_p,
\end{equation}
where, the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the 
vector separation between the two antennas. This signifies that the visibility 
($\widetilde{V}_p$) measured between a
pair of antennas ($p$) is obtained by the multiplying the sky brightness
$I(\hat{\mathbf{s}})$ by the antenna power response $B_p(\hat{\mathbf{s}})$ and
Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
coordinates, which are then sampled at the locations of the antenna spacings (or
baselines), namely, $\mathbf{u}_p$, and added to the receiver noise $n_p$.

This can be equivalently re-written as:
\begin{align}\label{eqn:software-holography}
  \widetilde{V}_p &= \int \widetilde{B}(\mathbf{u}^\prime-\mathbf{u}) \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u} + n_p,
\end{align}
where, $\widetilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
convolution with $\widetilde{B}(\mathbf{u})$ in $uv$-space. This is the software
holographic equivalent of a traditional FX correlator output.

@@

The intermediate step of gridding with the antenna power response can be 
expressed as a convolution of a data vector generated by gridding the electric 
fields directly with the antenna illumination pattern.
\begin{align}
\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v}) &= \left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]\right. \nonumber\\ 
&\quad\ast\left.\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
\end{align}

We can then use the multiplication-convolution theorem to move the convolution in 
Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in 
Equation 
\ref{eqn:dirty-image-FX}.
\begin{align}
  \mathbf{I}^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}^{-1}(\mathbf{a},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a})\,\right|^2\right\rangle_t. \label{eqn:dirty-image-MOFF}
\end{align}
The term inside the angular brackets before squaring has a very similar form as
that in equation~\ref{eqn:dirty-image-FX}. It signifies that the measured antenna
electric fields are weighted by the antenna noise, weighted and gridded by the
antenna aperture kernel, Fourier transformed and finally squared to obtain the
same image estimate that would have been obtained using equation~
\ref{eqn:dirty-image-FX}.
